---
title: "Step 0 Clean ExpiWell Data - Community W1"
author: "Hayleigh Armstrong & Jess Lougheed"
date: "11/02/2025"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

## Preliminary Steps

Load packages

```{r, warnings = FALSE, message = FALSE, include = FALSE}
library(tidyverse)
library(psych)
```

Read in the data:
```{r}
#File path on Hayleigh's Mac
df <- read.csv("/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_COMM_W1/Demographics_Community_numeric.csv", na.strings=c("","NA"))

# File path on Jess's Mac
#df <- read.csv("/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_COMM_W1/Demographics_Community_numeric.csv", na.strings=c("","NA"))

df <- as.data.frame(df)

```

Remove the responses of participants who were excluded from COMM W1 (based on Laura's table on Notion):
```{r}
df = filter(df, !(X.7 %in% c("67d3cf00c1ae9946c57f5245",
                             "680924677d7dd49024ad3129",
                             "68080b7008fd7252205f602c",
                             "681346515a3f3054a0754d08",
                             "680810ce08fd7252205fc83e",
                             "6807a56d08fd72522058a99b",
                             "681215895a3f3054a06a2114",
                             "681085385a3f305a05bad3e",
                             "681942ed07710cff0aa9fbb1",
                             "681945c307710cff0aaa3579",
                             "68198c2e07710cff0aace0f0",
                             "68198d8007710cff0aace763",
                             "681998e8307710cff0aacf1da",
                             "6819bbd307710cff0aad9216",
                             "68198f9307710cff0aacf6e5",
                             "681942ed07710cff0aa9fbb1",
                             "681945c307710cff0aaa3579",
                             "681a2ac507710cff0ab07a15",
                             "680b22be81fce634838cfb92",
                             "680b1abc81fce634838cdc67",
                             "6819a8c707710cff0aad540f",
                             "680b22be81fce634838cfb92",
                             "6807a56d08fd72522058a99b",
                             "680810ce08fd7252205fc83e",
                             "681a2ac507710cff0ab07a15",
                             "6812b3665a3f3054a071c382",
                             "680dfc745a3f3054a042bea0",
                             "680b1abc81fce634838cdc67",
                             "6819a8c707710cff0aad540f",
                             "68198e8307710cff0aacf1da",
                             "68198d8007710cff0aace763",
                             "68198c2e07710cff0aace0f0",
                             "681945c307710cff0aaa3579",
                             "681942ed07710cff0aa9fbb1",
                             "6819bbd307710cff0aad9216",
                             "6807f0bf08fd7252205dd059",
                             "680cdbe55a3f3054a037dcd5",
                             "680dfc745a3f3054a042bea0")))


```

```

Remove extra columns:
```{r}

cols_remove <- c("X.8", "X.9", "X.10", "X.11")
df <- df[, !(colnames(df) %in% cols_remove)]
```

## Save the data file

Save the data file. This is the file to read in to all COMM W1 ExpiWell variable preparation scripts.
```{r, eval = FALSE}
#On Hayleigh's Mac:
write.csv(df,'/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_COMM_W1/SEEDS COMM W1 Demographics ExpiWell Data Cleaned.csv')

# On Jess's Mac:
# write.csv(df,"/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_SONA_W1/SEEDS COMM W1 Demographics ExpiWell Data Cleaned.csv")
```
