---
title: "Step 1 BAI_SONA W1"
author: "Hayleigh Armstrong"
date: "06/10/2025"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

This script will process the data from the SONA W1 - Beck Anxiety Inventory.

## Preliminary Steps

Load packages:

```{r, warning = FALSE, message = FALSE, include = FALSE}
library(dplyr)
library(tidyverse)
library(Hmisc)
library(psych)
library(ggplot2)
library(ggcorrplot)
library(reshape2)
```

Read in the data:
```{r}
# Hayleigh's Mac: 
df <- read.csv("/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_SONA_W1/SEEDS SONA W1 Demographics ExpiWell Data Cleaned.csv", na.strings=c("","NA"))

# Jess's Mac:
#df <- read.csv("/Volumes/Emotion Dynamics Lab/PEARS/Data/SEEDS/PEARS_SEEDS_Demographics___Daily_Surveys_SONA_W1/SEEDS SONA W1 Demographics ExpiWell Data Cleaned.csv", na.strings=c("","NA"))

view(df)
```

Subset the data to just the items from the Beck Anxiety Inventory Scale and remove rows that have question headers:
```{r}
df_scale <- select(df [-c(1,2,3,4),], X.7, X.70:X.90)
View(df_scale)
```

Rename the items:
```{r}
colnames(df_scale)[colnames(df_scale)=="X.7"] <- "id"
colnames(df_scale)[colnames(df_scale)=="X.70"] <- "bai_1"
colnames(df_scale)[colnames(df_scale)=="X.71"] <- "bai_2"
colnames(df_scale)[colnames(df_scale)=="X.72"] <- "bai_3"
colnames(df_scale)[colnames(df_scale)=="X.73"] <- "bai_4"
colnames(df_scale)[colnames(df_scale)=="X.74"] <- "bai_5"
colnames(df_scale)[colnames(df_scale)=="X.75"] <- "bai_6"
colnames(df_scale)[colnames(df_scale)=="X.76"] <- "bai_7"
colnames(df_scale)[colnames(df_scale)=="X.77"] <- "bai_8"
colnames(df_scale)[colnames(df_scale)=="X.78"] <- "bai_9"
colnames(df_scale)[colnames(df_scale)=="X.79"] <- "bai_10"
colnames(df_scale)[colnames(df_scale)=="X.80"] <- "bai_11"
colnames(df_scale)[colnames(df_scale)=="X.81"] <- "bai_12"
colnames(df_scale)[colnames(df_scale)=="X.82"] <- "bai_13"
colnames(df_scale)[colnames(df_scale)=="X.83"] <- "bai_14"
colnames(df_scale)[colnames(df_scale)=="X.84"] <- "bai_15"
colnames(df_scale)[colnames(df_scale)=="X.85"] <- "bai_16"
colnames(df_scale)[colnames(df_scale)=="X.86"] <- "bai_17"
colnames(df_scale)[colnames(df_scale)=="X.87"] <- "bai_18"
colnames(df_scale)[colnames(df_scale)=="X.88"] <- "bai_19"
colnames(df_scale)[colnames(df_scale)=="X.89"] <- "bai_20"
colnames(df_scale)[colnames(df_scale)=="X.90"] <- "bai_21"
```

Convert the "prefer not to respond" option to "NA":
```{r}
df_scale[df_scale == 4] <- NA

```

Variables are pulled in as factors and we want to make them numeric values: 
```{r}
df_scale[,paste0("bai_",1:21)] <- lapply(df_scale[,paste0("bai_",1:21)], as.numeric)
str(df_scale)
```

## Calculate the scale

### Calculate the standard mean and sum

Sum and mean scores of Beck Anxiety Inventory Scale: 
```{r}

df_scale$bai_sum <- apply(df_scale[,paste0("bai_",1:21)], 1, sum)
df_scale$bai_mean <- apply(df_scale[,paste0("bai_",1:21)], 1, mean)
df_scale$bai_mean <- round(df_scale$bai_mean, digits = 2)

```

Label the items:
```{r}
label(df_scale[["bai_1"]]) <- "Numbness and tingling"
label(df_scale[["bai_2"]]) <- "Feeling hot"
label(df_scale[["bai_3"]]) <- "Wobbliness in legs"
label(df_scale[["bai_4"]]) <- "Unable to relax"
label(df_scale[["bai_5"]]) <- "Fear of worst happening"
label(df_scale[["bai_6"]]) <- "Dizzy or lightheaded"
label(df_scale[["bai_7"]]) <- "Heart pounding or racing"
label(df_scale[["bai_8"]]) <- "Unsteady"
label(df_scale[["bai_9"]]) <- "Terrified"
label(df_scale[["bai_10"]]) <- "Nervous"
label(df_scale[["bai_11"]]) <- "Feelings of choking"
label(df_scale[["bai_12"]]) <- "Hands trembling"
label(df_scale[["bai_13"]]) <- "Shaky"
label(df_scale[["bai_14"]]) <- "Fear of losing control"
label(df_scale[["bai_15"]]) <- "Difficulty breathing"
label(df_scale[["bai_16"]]) <- "Fear of dying"
label(df_scale[["bai_17"]]) <- "Scared"
label(df_scale[["bai_18"]]) <- "Abdominal discomfort"
label(df_scale[["bai_19"]]) <- "Faint"
label(df_scale[["bai_20"]]) <- "Face flushed"
label(df_scale[["bai_21"]]) <- "Sweating (not due to heat)"

label(df_scale[["bai_sum"]]) <- "sum of bai items"
label(df_scale[["bai_mean"]]) <- "mean of bai items"

# Look at the data
head(df_scale)
```

## Save the data 

Save the data file:
```{r, eval = FALSE}
# Enya's Mac:
write.csv(df_scale,'/Volumes/Emotion Dynamics Lab/DREAMS (UGS2)/Data/Part 1/Term 1 Data/Beck Anxiety Inventory-Part_1.csv')

# Jess's Mac:
#write.csv(df_scale,'/Volumes/research/Emotion Dynamics Lab/DREAMS (UGS2)/Data/Part 1/Term 1 Data/Beck Anxiety Inventory-Part_1.csv')
```

## Examine the Data

### Reliability
```{r, warning = FALSE}
items <- c ("bai_1",  "bai_2" , "bai_3" , "bai_4" , "bai_5" , "bai_6" , "bai_7" , "bai_8" , "bai_9",  "bai_10" , "bai_11" , "bai_12" , "bai_13" , "bai_14" , "bai_15" , "bai_16" , "bai_17",  "bai_18" , "bai_19" , "bai_20" , "bai_21")
alpha(df_scale[,items])
```

### Descriptive statistics
```{r}
describe(df_scale$bai_sum)
describe(df_scale$bai_mean)
```

### Histogram
```{r, warning = FALSE, message = FALSE}
qplot(df_scale$bai_mean, geom="histogram") 
```